Left-Linear Rewriting in Adhesive Categories

### 2024-07-02 concur/final.tex
aggiustatina a directory e ack

\documentclass[a4paper,UKenglish,cleveref,pdftex,thm-restate,numberwithinsect]{lipics}
\newif\ifreport
\reporttrue 
%\reportfalse

@@

\usepackage{pgf}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri,fit,positioning}
\tikzstyle{node}=[circle, draw=black, minimum size=1mm, inner sep=1.5pt]
\tikzstyle{trans}=[font=\scriptsize]
\tikzstyle{lab}=[font=\small]
\tikzset{
  coloredge/.style={
        ->,
        color=red
        %densely dotted
    }
}

@@

@@

\author{Paolo Baldan} 
{Department of Mathematics, University of Padova, Italy}
{baldan@math.unipd.it}{}{}

\author{Davide Castelnovo}
{Department of Mathematics, University of Padova, Italy}
{davide.castelnovo@math.unipd.it}{}{}
\author{Andrea Corradini}
{Department of Computer Science, University of Pisa, Italy}
 {andrea.corradini@unipi.it}{}{}

\author{Fabio Gadducci}
{Department of Computer Science, University of Pisa, Italy}
{fabio.gadducci@unipi.it}{}{}

\authorrunning{P.~Baldan, D.~Castelnovo, A.~Corradini, F.~Gadducci}
%
\Copyright{Paolo Baldan, Davide Castelnovo, Andrea Corradini and Fabio Gadducci}
%
\ccsdesc[500]{Theory of Computation~Models of computation}
%
\keywords{
Adhesive categories, double-pushout rewriting, left-linear rules, switch equivalence, local Church-Rosser property.
}

@@

Here we perform this exploration at the level of $\mathcal{M}$-adhesive categories, 
which encompass a large family of DPO-based formalisms adopted by researchers and 
practitioners alike.

@@

\bibliography{bibliog.bib}

>>>>

\documentclass[a4paper,UKenglish,cleveref,pdftex,thm-restate,numberwithinsect]{lipics-v2021}
\newif\ifreport
\reporttrue 
%\reportfalse

@@

\usepackage{pgf}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri,fit,positioning}
\tikzstyle{node}=[circle, draw=black, minimum size=1mm, inner sep=1.5pt, font=\tiny]

\tikzstyle{trans}=[font=\scriptsize]
\tikzstyle{lab}=[font=\small]
\tikzset{
  coloredge/.style={
        ->,
        color=red
        %densely dotted
    }
}

@@

\EventEditors{Rupak Majumdar and Alexandra Silva}
\EventNoEds{2}
\EventLongTitle{35th International Conference on Concurrency Theory (CONCUR 2024)}
\EventShortTitle{CONCUR 2024}
\EventAcronym{CONCUR}
\EventYear{2024}
\EventDate{Septembre 9--13, 2024}
\EventLocation{ }
\EventLogo{ }
\SeriesVolume{xxx}
\ArticleNo{xxx}

@@

\author{Paolo Baldan} 
{Department of Mathematics, University of Padova, Italy}
{baldan@math.unipd.it}
{}
{xxx}

\author{Davide Castelnovo}
{Department of Mathematics, University of Padova, Italy}
{davide.castelnovo@math.unipd.it}
{}
{yyy}

\author{Andrea Corradini}
{Department of Computer Science, University of Pisa, Italy}
 {andrea.corradini@unipi.it}
 {}
 {MUR PRIN 20228KXFN2 ``STENDHAL'' and U. Pisa PRA 2022\_99 ``FM4HD''}

\author{Fabio Gadducci}
{Department of Computer Science, University of Pisa, Italy}
{fabio.gadducci@unipi.it}
{0000-0002-9123-9089}
 {MUR PRIN 20228KXFN2 ``STENDHAL'' and U. Pisa PRA 2022\_99 ``FM4HD''}

\authorrunning{P.~Baldan, D.~Castelnovo, A.~Corradini, F.~Gadducci}
%
\Copyright{Paolo Baldan, Davide Castelnovo, Andrea Corradini, and Fabio Gadducci}
%
\ccsdesc[500]{Theory of Computation~Models of computation}
%
\keywords{
Adhesive categories, double-pushout rewriting, left-linear rules, switch equivalence, local Church-Rosser property.
}

@@

\begin{figure}
  \input{Figure/coffee-rules.tex}

  \caption{Rewirting rules for the coffee example.}
  \label{fi:coffee-rules}
\end{figure}

\begin{figure}
  \input{Figure/coffee-rewriting.tex}

  \caption{A sequence of rewriting steps in .}
  \label{fi:coffee-rewriting}
\end{figure}

Just to get some basic intuition about left-linear rewriting systems, their capabilities and the problems which can arise, assume we want to model a situation in which a coffee comes with a bag of sugar, a portion of milk and bag of cocoa. The situation is represented by the graph $G_0$ in Fig.~\ref{fi:coffee-rewriting}: the loop $\mathit{cf}$ is the coffee and the three edges $m$, $s$ and $c$  represent the milk, sugar and cocoa, respectively. Adding one of these elements, call it $x$, in the coffee is modeled as merging the corresponding node with the coffee node so that a self-loop $x$ is created.

The available rules are in Fig.~\ref{fi:coffee-rules}. Rules $\rho_m$ models the addition of milk and sugar, while $\rho_c$ adds cocoa and sugar to the coffee. Once the sugar has been added, the coffee can be stirred and then it is ready to drink. This is represented by rule $\rho_r$ which check for the presence of sugar in the coffee, and stirs it.

A possible rewriting sequence is depicted in Fig.~\ref{fi:coffee-rewriting}, applying $\rho_m$, $\rho_c$ and $\rho_r$ in sequence, thus producing a coffee with milk, sugar and cocoa, ready for drinking. A first interesting observation is that after applying rule $\rho_m$ to $G_0$, thus producing graph $G_1$, we can still apply rule $\rho_c$, with a non-injective match. Intuitively, the two rules do not interfere. Rather, $\rho_c$ applied after $\rho_m$ will have no effect on the sugar edge (intuitively, because sugar has been already added). Indeed, the rules could be also be applied in reverse order. Should they be considered independent? Moreover, since rule $\rho_r$ just require the presence of sugar, it could be anticipated, and applied after $r_m$. This might lead to think that $\rho_c$ and $\rho_m$ are independent. However, this is not completely clear, since in absence of $\rho_m$, the application of $\rho_c$ would be essential to enable $\rho_r$. ``There is great chaos under heaven, [hence] the situation is excellent!''

In this paper we perform a systematic exploration of these phenomena at the
% Here we perform this exploration at the
level of $\mathcal{M}$-adhesive categories,
which encompass a large family of DPO-based formalisms adopted by researchers and 
practitioners alike.

@@

\bibstyle{plainurl}
\bibliography{bibliog.bib}

### 2024-07-02 concur/submitted/fusione.tex
aggiustatina a directory e ack

\documentclass[a4paper,UKenglish,cleveref,pdftex,thm-restate,numberwithinsect]{lipics-v2021}
\newif\ifreport
\reporttrue 
%\reportfalse

@@

\EventEditors{Rupak Majumdar and Alexandra Silva}
\EventNoEds{2}
\EventLongTitle{35th International Conference on Concurrency Theory (CONCUR 2024)}
\EventShortTitle{CONCUR 2024}
\EventAcronym{CONCUR}
\EventYear{2024}
\EventDate{Septembre 9--13, 2024}
\EventLocation{ }
\EventLogo{ }
\SeriesVolume{xxx}
\ArticleNo{xxx}

@@

\author{Paolo Baldan} 
{Department of Mathematics, University of Padova, Italy}
{baldan@math.unipd.it}
{}
{xxx}

\author{Davide Castelnovo}
{Department of Mathematics, University of Padova, Italy}
{davide.castelnovo@math.unipd.it}
{}
{yyy}

\author{Andrea Corradini}
{Department of Computer Science, University of Pisa, Italy}
 {andrea.corradini@unipi.it}
 {}
 {MUR PRIN 20228KXFN2 ``STENDHAL'' and U. Pisa PRA 2022\_99 ``FM4HD''}

\author{Fabio Gadducci}
{Department of Computer Science, University of Pisa, Italy}
{fabio.gadducci@unipi.it}
{0000-0002-9123-9089}
 {MUR PRIN 20228KXFN2 ``STENDHAL'' and U. Pisa PRA 2022\_99 ``FM4HD''}

\authorrunning{P.~Baldan, D.~Castelnovo, A.~Corradini, F.~Gadducci}
%
\Copyright{Paolo Baldan, Davide Castelnovo, Andrea Corradini, and Fabio Gadducci}
%
\ccsdesc[500]{Theory of Computation~Models of computation}
%
\keywords{
Adhesive categories, double-pushout rewriting, left-linear rules, switch equivalence, local Church-Rosser property.
}

@@

\bibstyle{plainurl}
\bibliography{bibliog.bib}

>>>>

\documentclass[a4paper,UKenglish,cleveref,pdftex,thm-restate,numberwithinsect]{lipics}
\newif\ifreport
\reporttrue 
%\reportfalse

@@

@@

\author{Paolo Baldan} 
{Department of Mathematics, University of Padova, Italy}
{baldan@math.unipd.it}{}{}

\author{Davide Castelnovo}
{Department of Mathematics, University of Padova, Italy}
{davide.castelnovo@math.unipd.it}{}{}
\author{Andrea Corradini}
{Department of Computer Science, University of Pisa, Italy}
 {andrea.corradini@unipi.it}{}{}

\author{Fabio Gadducci}
{Department of Computer Science, University of Pisa, Italy}
{fabio.gadducci@unipi.it}{}{}

\authorrunning{P.~Baldan, D.~Castelnovo, A.~Corradini, F.~Gadducci}
%
\Copyright{Paolo Baldan, Davide Castelnovo, Andrea Corradini and Fabio Gadducci}
%
\ccsdesc[500]{Theory of Computation~Models of computation}
%
\keywords{
Adhesive categories, double-pushout rewriting, left-linear rules, switch equivalence, local Church-Rosser property.
}

@@

\bibliography{bibliog.bib}
